Finite element analysis ofmulti-dimensional and simplifiedmodels for beams and plates

### 2023-12-03 files/Overview.tex
overview

\section{Achievements}

@@

In the fourth chapter, we look at our simplest model, the Timoshenko beam model. The first result discusses an article that provides a method to determine the exact eigenvalues and eigenfunctions for the Timoshenko beam model. We then apply this to a cantilever beam model and a free-free beam model to obtain the eigenvalues and plot the mode shapes. We then look at an article that compares a cantilever Timoshenko beam model to a two-dimensional beam model. We discuss the results of the article and replicate the results with more significant digits. We also include some more results not included in the article, that are used later in Chapter 6. We then look at another article that compares the Timoshenko beam theory to results obtained from a measured beam. The authors suspended a beam on both ends an vibrated the beam until then endpoints could be considered free. They then swept through different frequencies and recorded the natural frequencies of the beam. These results are then compared to the theoretical results of a free-free Timoshenko beam. We add to the article by giving the model problems for the beams and discuss the results.

>>>>

\section{Achievements and contributions}

@@

We then investigate the validity of the Timoshenko beam theory. This is presented in an article where the cantilever Timoshenko beam theory is compared to a cantilever two-dimensional beam model. We apply the method use by the authors to plot the mode shapes. These mode shapes are then compared and the eigenvalues are matched up. This is required since there are eigenvalues for the two-dimensional model that are not related to beam type eigenvalues. The authors make a distinction by referring to the eigenvalues as beam-type eigenvalues and non-beam type eigenvalues. For the Timoshenko beam model, the same method and computer program used in this chapter was used. For the two-dimensional model, the Finite Element Method is used. This is explained more in detail in chapter 5, and in chapter 6. The results of the article is replicated, with five significant digits of accuracy (the article use three significant digits of accuracy). Also more cases are presented to expand on the article as well as in preparation for the results in chapter 6.

The final result of this chapter looks at a study that compares empirical results from a real world study on the vibrations of a beam to the theoretical results of the Timoshenko beam model. We add to the article by including the model problems.

The fifth chapter is preparation for the comparisons of the models in this dissertation. In chapter 4, we discussed a method to obtain to obtain the exact solutions for the eigenvalue problem for the Timoshenko beam model. For the other models a different approach is required. For each of the models, the Galerkin Approximation is derived. The reference configuration is the divided into discrete elements. We then apply piece-wise Hermite cubic functions and use the Finite Element Method to obtain a initial value problem for each of the models consisting of a system of ordinary differential equations. An eigenvalue problem is then derived from this IVP. Using a computer program, this eigenvalue problem can be solved and the eigenvalues and eigenvectors can be calculated. An investigation into the rate of convergence was conducted to make sure that the results are accurate enough.
